### 2015-09-11 vmcai-2016-tex/approach.tex
All sections were edited to remove first-person voice, e.g. we, our, etc.

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
reference superscripts $s$ and $i$ indicate the partition containing
the reference.  In~\figref{fig:initHeap}, $\cfgnt{r}_0^s$ represents a
stack reference for the $\cfgt{this}$ variable which has two fields
$x$ and $y$ of the same type. Note that when no constraint is
specified for a location, there is an implicit $\mathit{true}$
constraint. For example, $\cfgnt{r}_0^s$ points to $\cfgnt{l}_0$ on
the constraint $\mathit{true}$. The fields $x$ and $y$ point to the
uninitialized reference $\cfgnt{r}_\mathit{un}$. The graph
in~\figref{fig:initHeap}(b) represents the symbolic heap after the
initialization of the $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$ field
while the graph in~\figref{fig:initHeap}(c) represents the symbolic
heap after the initialization of the
$\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$ field following the
initialization of $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
the initialization $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$.

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other when we write to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location. The field-access rewrite rule
in~\figref{fig:fHeap} dereferences a field of type $\cfgnt{C}$ and
uses \figref{fig:symInit}, to get a new symbolic heap that is
initialized on the field $\cfgnt{f}$. The symbolic heap is further
modified to include a new stack reference pointing to the guarded value set
(possible values of the field) returned during the dereferencing; the
new stack reference is the return value from the operation.

@@

The rewrite rule to compare two references in the symbolic heap is
shown in~\figref{fig:eqs}. The equals references-true rewrite rule
returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$
\emph{can} be equal. Intutively, $\Phi_\alpha$ contains all
constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may point to
the same location in the symbolic heap. The second set, $\Phi_0$,
contains constraints under which the reference $\cfgnt{r}_0$ points to
corresponding locations such that the reference $\cfgnt{r}_1$
\emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and
$\cfgnt{r}_0$ \emph{does not} point to those locations under any
constraint. We use the three sets of constraints to create a new path condition $\phi_g^\prime$
as an update of the current path condition $\phi_g$. We add to $\phi_g$ the disjunction of the
constraints in $\Phi_\alpha$ to indicate that if any of the
constraints are satisfiable, then references $\cfgnt{r}_0$ and
$\cfgnt{r}_1$ can be equal. Furthermore, we conjoin to the global heap
constraint $\phi_g^\prime$, the conjunctions of negations to the
constraints in $\Phi_0$ and $\Phi_1$. This indicates for locations
that are not common to the references, the negations of their
constraints are satisfiable. The rule does not complete (i.e., is not
feasible) if the new global constraint is not satisfied on any
aliasing assignment. In such a case, the $\cfgt{true}$ outcome is not
possible on any symbolic heap. Before the rewrite rule returns
$\cfgt{true}$ we check the satisfiability of the updated global heap
constraint. The reference-false is the logical dual of the rule.

Consider the example in~\figref{fig:initHeap}(c). In order to to check
if $\cfgnt{r}_1^i$ and $\cfgnt{r}_2^i$ are equal we first get the 
guarded value set associated with each of the references:
\[
\cfgnt{L}(\cfgnt{r}_1^i) = \{ (\phi_{1a}\; \cfgnt{l}_\mathit{null})\; (\phi_{1b}\; \cfgnt{l}_1) \} 
\]
\[
\cfgnt{L}(\cfgnt{r}_2^i) = \{ (\phi_{2a}\; \cfgnt{l}_\mathit{null})\; (\phi_{2b}\; \cfgnt{l}_2)\; (\phi_{2c}\; \cfgnt{l}_1) \} \\
\]
\noindent{The three constraint sets are:} 
\[
\Phi_\alpha = \{ (\phi_{1a}\; \wedge\; \phi_{2a} ) (\phi_{1b}\; \wedge\; \phi_{2c} ) \}\;
\Phi_0 = \{ \}\; \Phi_1 = \{ \phi_{2b}\} \\
\]
\noindent{Finally the global constraint is} 
\[
\phi^\prime = \mathit{true} \wedge [ (\phi_{1a}\; \wedge\; \phi_{2a} )\vee (\phi_{1b}\; \wedge\; \phi_{2c} ) ] \wedge \neg\phi_{2b} 
\]

>>>>

\figref{fig:initHeap} illustrates the initialization process. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
reference superscripts $s$ and $i$ indicate the partition containing
the reference.  In~\figref{fig:initHeap}, $\cfgnt{r}_0^s$ represents a
stack reference for the $\cfgt{this}$ variable which has two fields
$x$ and $y$ of the same type. Note that when no constraint is
specified for a location, there is an implicit $\mathit{true}$
constraint. For example, $\cfgnt{r}_0^s$ points to $\cfgnt{l}_0$ on
the constraint $\mathit{true}$. The fields $x$ and $y$ point to the
uninitialized reference $\cfgnt{r}_\mathit{un}$. The graph
in~\figref{fig:initHeap}(b) represents the symbolic heap after the
initialization of the $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$ field
while the graph in~\figref{fig:initHeap}(c) represents the symbolic
heap after the initialization of the
$\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$ field following the
initialization of $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{x}\rp$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
the initialization $\lp\cfgt{this}\ \cfgt{\$}\ \cfgnt{y}\rp$.

@@

There are two rewrite rules in~\figref{fig:fHeap}(a), one for reading
the value of a field (field-access) and the other for writing to a
field (field-write). Both rules first check that the operations can be
performed on a non-null location. The field-access rewrite rule
in~\figref{fig:fHeap} dereferences a field of type $\cfgnt{C}$ and
uses \figref{fig:symInit}, to get a new symbolic heap that is
initialized on the field $\cfgnt{f}$. The symbolic heap is further
modified to include a new stack reference pointing to the guarded value set
(possible values of the field) returned during the dereferencing; the
new stack reference is the return value from the operation.

@@

The rewrite rule to compare two references in the symbolic heap is
shown in~\figref{fig:eqs}. The equals references-true rewrite rule
returns true when two references $\cfgnt{r}_0$ and $\cfgnt{r}_1$
\emph{can} be equal. Intutively, $\Phi_\alpha$ contains all
constraints under which $\cfgnt{r}_0$ and $\cfgnt{r}_1$ may point to
the same location in the symbolic heap. The second set, $\Phi_0$,
contains constraints under which the reference $\cfgnt{r}_0$ points to
corresponding locations such that the reference $\cfgnt{r}_1$
\emph{does not} point to those locations under any
constraint. Finally, the set, $\Phi_1$, contains constraints under
which $\cfgnt{r}_1$ points to corresponding locations and
$\cfgnt{r}_0$ \emph{does not} point to those locations under any
constraint. The three sets of constraints are used to create a new path condition $\phi_g^\prime$
as an update of the current path condition $\phi_g$.  The update is accomplished by first taking the disjunction of the
constraints in $\Phi_\alpha$ to indicate that if any of the
constraints are satisfiable, then references $\cfgnt{r}_0$ and
$\cfgnt{r}_1$ can be equal. This disjunction is then conjoined with $\phi_g$ to form $\phi_g^\prime$. Furthermore, the conjunctions of negations to the
constraints in $\Phi_0$ and $\Phi_1$ is conjoined with $\phi_g^\prime$. This indicates for locations
that are not common to the references, the negations of their
constraints are satisfiable. The rule does not complete (i.e., is not
feasible) if the new global constraint is not satisfied on any
aliasing assignment. In such a case, the $\cfgt{true}$ outcome is not
possible on any symbolic heap. Before the rewrite rule returns
$\cfgt{true}$, it verifies the satisfiability of the updated global heap
constraint. The reference-false is the logical dual of the rule.

Consider the example in~\figref{fig:initHeap}(c). In order to compare
$\cfgnt{r}_1^i$ and $\cfgnt{r}_2^i$, the Equals rule gets the 
guarded value set associated with each of the references:
\[
\cfgnt{L}(\cfgnt{r}_1^i) = \{ (\phi_{1a}\; \cfgnt{l}_\mathit{null})\; (\phi_{1b}\; \cfgnt{l}_1) \} 
\]
\[
\cfgnt{L}(\cfgnt{r}_2^i) = \{ (\phi_{2a}\; \cfgnt{l}_\mathit{null})\; (\phi_{2b}\; \cfgnt{l}_2)\; (\phi_{2c}\; \cfgnt{l}_1) \} \\
\]
\noindent{The three constraint sets are:} 
\[
\Phi_\alpha = \{ (\phi_{1a}\; \wedge\; \phi_{2a} ) (\phi_{1b}\; \wedge\; \phi_{2c} ) \}\;
\Phi_0 = \{ \}\; \Phi_1 = \{ \phi_{2b}\} \\
\]
\noindent{Finally the global constraint is} 
\[
\phi^\prime = \mathit{true} \wedge [ (\phi_{1a}\; \wedge\; \phi_{2a} )\vee (\phi_{1b}\; \wedge\; \phi_{2c} ) ] \wedge \neg\phi_{2b} 
\]

### 2015-09-11 vmcai-2016-tex/conclusion.tex
All sections were edited to remove first-person voice, e.g. we, our, etc.

In this work, we present a
heap representation for symbolic execution that maintains exactness,
represents arbitrary heap structures without restrictions on aliasing
relationships, and supports dynamic on-the-fly initialization of heap
structures. This unique combination of capabilities enables the
symbolic heap algorithm to outperform state-of-the-art symbolic
execution techniques, and in some cases, the performance gains are
considerable. We believe that performance gains in symbolic execution
can naturally benefit a variety of analyses based on symbolic
execution, for example, verification of properties, especially those
related to the heap, test case generation, program evolution
techniques such as directed incremental symbolic
execution~\cite{person:pldi2011}, and other back-end analyses. In
future work we plan to evaluate the impact of analyses that rely on
symbolic execution.

>>>>

This work presents a
heap representation for symbolic execution that maintains exactness,
represents arbitrary heap structures without restrictions on aliasing
relationships, and supports dynamic on-the-fly initialization of heap
structures. This unique combination of capabilities enables the
symbolic heap algorithm to outperform state-of-the-art symbolic
execution techniques, and in some cases, the performance gains are
considerable. The performance gains in symbolic execution
may naturally benefit a variety of analyses based on symbolic
execution, for example, verification of properties, especially those
related to the heap, test case generation, program evolution
techniques such as directed incremental symbolic
execution~\cite{person:pldi2011}, and other back-end analyses. Potential future work includes evaluating the impact of analyses that rely on
symbolic execution.

### 2015-09-11 vmcai-2016-tex/evaluation.tex
All sections were edited to remove first-person voice, e.g. we, our, etc.

To resolve this question, we implemented a prototype of the \symtxt{} 
algorithm in Symbolic PathFinder (SPF)~\cite{DBLP:journals/ase/PasareanuVBGMR13} 
and evaluate its efficacy on Java data structure artifacts.
%SPF includes an implementation of~\gsetxt{} with lazy initialization.
%In recent work, we implemented the  Lazier and Lazier\#
%algorithms in SPF~\cite{Hillery:2014}.
%
%The prototype of our~\symtxt{} algorithm is implemented as an
%extension to SPF.  
In addition to the operations presented in
this paper, the implementation contains support for operations over
integers, calculating per-path preconditions 
and postconditions, as well as path-coverage-optimal sets of 
test input heaps. 
%A standalone version of the system, independent
%from JPF is also under development~\cite{heapsolver2015}.

@@

The results of the experiments are presented in Table~\ref{tab:results}. 
Each row reports the results for the specified $k$-bound 
for each artifact evaluated. The columns show the total
time in seconds, states explored, and paths generated for
each algorithm. The headings \gsetxt{}, L\#, and SH correspond
to the Generalized Symbolic Execution, Lazier\#, and \symtxt{}
algorithms, respectively. Note that we omit the results for
the Lazier algorithm because they are nearly identical to the Lazier\#
results for these artifacts. A table entry of  '-' indicates the analysis
exceeded the allotted time bound of 1 hour.

>>>>

To resolve this question, a prototype of the \symtxt{} 
algorithm was developed in the Symbolic PathFinder (SPF)~\cite{DBLP:journals/ase/PasareanuVBGMR13} framework
to evaluate its efficacy on Java data structure artifacts.
%SPF includes an implementation of~\gsetxt{} with lazy initialization.
%In recent work, we implemented the  Lazier and Lazier\#
%algorithms in SPF~\cite{Hillery:2014}.
%
%The prototype of our~\symtxt{} algorithm is implemented as an
%extension to SPF.  
In addition to the operations presented in
this paper, the implementation contains support for operations over
integers, calculating per-path preconditions 
and postconditions, as well as path-coverage-optimal sets of 
test input heaps. 
%A standalone version of the system, independent
%from JPF is also under development~\cite{heapsolver2015}.

@@

The results of the experiments are presented in Table~\ref{tab:results}. 
Each row reports the results for the specified $k$-bound 
for each artifact evaluated. The columns show the total
time in seconds, states explored, and paths generated for
each algorithm. The headings \gsetxt{}, L\#, and SH correspond
to the Generalized Symbolic Execution, Lazier\#, and \symtxt{}
algorithms, respectively. The results for
the Lazier algorithm are omitted because they are nearly identical to the Lazier\#
results for these artifacts. A table entry of  '-' indicates the analysis
exceeded the allotted time bound of 1 hour.

### 2015-09-11 vmcai-2016-tex/lazy.tex
All sections were edited to remove first-person voice, e.g. we, our, etc.

The~\gsetxt{} initialization rules preserve several important properties of the symbolic heap. In the heap generated by~\gsetst{}, each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals}.
  Only input heap references appear in constraints to express potential aliasing.
  Auxiliary literals and stack references are used to define additional structure in the heap but only stack references can appear in environments, expressions, or continuations. Changes to $R$ are monotonically increasing, eliminating any need for constraint rewriting as the program state evolves.
A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

>>>>

The~\gsetxt{} initialization rules preserve several important properties of the symbolic heap. In the heap generated by~\gsetxt{}, each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals}.
  Only input heap references appear in constraints to express potential aliasing.
  Auxiliary literals and stack references are used to define additional structure in the heap but only stack references can appear in environments, expressions, or continuations. Changes to $R$ are monotonically increasing, eliminating any need for constraint rewriting as the program state evolves.
A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

### 2015-09-11 vmcai-2016-tex/rw.tex
All sections were edited to remove first-person voice, e.g. we, our, etc.

Recently, value sets have been adapted for use in symbolic execution, for the purposes of state merging ~\cite{Sen:2014,Torlak:2014} and invariant detection~\cite{Ferrara:2014}. While these methods demonstrate the utility of value sets in combination with symbolic execution, none of these techniques address dereferencing of symbolic input references. Since dereferencing is a fundamental problem treated in this paper, we believe these other works to be both orthogonal and complimentary to our own.
%
%\subsection{Other Heap Representations in Symbolic Execution}
%Need to talk about GSE, DSE, SAGE, PEX
%GSE
Several projects have generalized the core idea of symbolic execution,
enabling it to be applied to programs with more general types,
including references and
arrays~\cite{GSE03,KiasanKunit,Cadar:2008,Rosner:2015}. However, these
generalized symbolic execution (\gsetxt{}) techniques work by branching over
multiple copies of the system state during dereferencing operations,
exacerbating path explosion.

%DSE
Dynamic symbolic execution (\dsetxt{}) techniques may use
concretization when reasoning about references~\cite{Godefroid:2005,Sen:2005,Godefroid:POPL07,Tillmann:2008}, enabling much greater program
coverage than classical symbolic execution alone. In particular, PEX~\cite{Tillmann:2008}
has demonstrated a high degree of utility in a wide variety of 
practical verification scenarios. Sometimes,
however, as demonstrated in~\cite{Elkarablieh:2009},~\dsetxt{}
under-approximates the possible heaps configurations when it
concretizes symbolic values. The SAGE~\dsetxt{}
engine~\cite{Elkarablieh:2009} includes more sophisticated methods for
reasoning symbolically about references, and can successfully handle
complicated examples. However, SAGE uses a
theory of fixed-length arrays to model the heap, and is thus incomplete
for programs with input-dependent memory allocation. We believe that the 
techniques presented in this work can be used to extend practical tools 
like PEX for applications requiring formal correctness guarantees.

%
%\subsection{Formal Results on Symbolic Heaps}
The symbolic heap used by SAGE~\cite{Elkarablieh:2009} was shown to be sound and complete for programs with input-independent memory allocation. Our method is sound and complete for all states reached by symbolic execution.

Recently, several separation logic solvers have been proven to be sound and complete for heaps with linked lists, trees~\cite{Piskac:2014}, or data structures satisfying user-supplied invariants~\cite{Brotherston:2014}. In contrast, our method is sound and complete without restriction on data structures, and requires no explicit invariants, but our result is restricted to the properties provable by symbolic execution.

>>>>

Recently, value sets have been adapted for use in symbolic execution, for the purposes of state merging ~\cite{Sen:2014,Torlak:2014} and invariant detection~\cite{Ferrara:2014}. While these methods demonstrate the utility of value sets in combination with symbolic execution, none of these techniques address dereferencing of symbolic input references. Since dereferencing is a fundamental problem treated in this paper, these other works may be considered both orthogonal and complimentary to~\symtxt{}.
%
%\subsection{Other Heap Representations in Symbolic Execution}
%Need to talk about GSE, DSE, SAGE, PEX
%GSE
Several projects have generalized the core idea of symbolic execution,
enabling it to be applied to programs with more general types,
including references and
arrays~\cite{GSE03,KiasanKunit,Cadar:2008,Rosner:2015}. However, these
generalized symbolic execution (\gsetxt{}) techniques work by branching over
multiple copies of the system state during dereferencing operations,
exacerbating path explosion.

%DSE
Dynamic symbolic execution (\dsetxt{}) techniques may use
concretization when reasoning about references~\cite{Godefroid:2005,Sen:2005,Godefroid:POPL07,Tillmann:2008}, enabling much greater program
coverage than classical symbolic execution alone. In particular, PEX~\cite{Tillmann:2008}
has demonstrated a high degree of utility in a wide variety of 
practical verification scenarios. Sometimes,
however, as demonstrated in~\cite{Elkarablieh:2009},~\dsetxt{}
under-approximates the possible heaps configurations when it
concretizes symbolic values. The SAGE~\dsetxt{}
engine~\cite{Elkarablieh:2009} includes more sophisticated methods for
reasoning symbolically about references, and can successfully handle
complicated examples. However, SAGE uses a
theory of fixed-length arrays to model the heap, and is thus incomplete
for programs with input-dependent memory allocation. It is hoped that the 
techniques presented in this work can be used to extend practical tools 
like PEX for applications requiring formal correctness guarantees.

%
%\subsection{Formal Results on Symbolic Heaps}
The symbolic heap used by SAGE~\cite{Elkarablieh:2009} was shown to be sound and complete for programs with input-independent memory allocation.~\symtxt{} is sound and complete for all states reached by symbolic execution.

Recently, several separation logic solvers have been proven to be sound and complete for heaps with linked lists, trees~\cite{Piskac:2014}, or data structures satisfying user-supplied invariants~\cite{Brotherston:2014}. In contrast,~\symtxt{} is sound and complete without restriction on data structures, and requires no explicit invariants, but this result is restricted to the properties provable by symbolic execution.
